---
title: "Data Science Project"
author: "Gege Gui"
date: "9/11/2017"
output: pdf_document
---

# Introduction:

PLoS (Public Library of Science) is a nonprofit Open Access publisher, innovator and advocacy organization with a mission to accelerate progress in science and medicine by leading a transformation in research communication. The data analysis project is to find out the most common statistical techniques in all published PLoS papers and the trends over the last 10-15 years.

# Data:

Follow the documentation from PLoS API website <http://api.plos.org/solr/examples/> to download statistics-related paper. By indicating "statistic" in abstract part, the results show that there are 12329 papers. 

```{r}
library("rplos")
# Get only full article DOIs
id = searchplos("abstract:statistics", fl='id, publication_date', fq='doc_type:full', sort = 'publication_date desc', start=0)
```

Specify the "limit" to be 12329, then the data part for "id" contains the ID and publication date of the paper of our interest.

```{r}
id = searchplos("abstract:statistics", fl='id, publication_date', fq='doc_type:full', sort = 'publication_date desc', start=0, limit = id$meta[1, 1])
```

To use regular expression for further search, we create a dictionary for common statistical methods. We use "The Elements of Statistical Learning - Data Mining, Inference, and Prediction" second edition, by Trevor Hastie, Robert Tibshirani, Jerome Friedman. 

```{r}
dic = c("linear regression","Lasso","ridge regression","linear discriminant analysis (LDA)","bootstrap","maximum likelihood","EM algorithm","MCMC","neural network","support vector machine (SVM)","clustering","principal component","independent component","random forest","graphical model")
```

```{r}
# create dictionary and return abstract
plosabstract(q = 'regression', fl='id,title', limit = 5)
plosabstract(q = 'machine learning', fl='id,title', limit = 5)
plosabstract(q = 'neural network', fl='id,title', limit = 5)
```

```{r}
# select paper with key words in method part
out <- highplos(q='regression', hl.fl = 'Materials and Methods', rows=130)
highbrow(out)
```

```{r}
# Visualize word use across articles
plosword(list('regression', 'neural network'), vis = 'TRUE')

# plot through time
pl_t = plot_throughtime(terms = c('regression', 'neural network'), limit = 59)
```


# Related paper

Topics over Time: A Non-Markov Continuous-Time Model of Topical Trends

Use graphical model and Gibbs sampling to find the topic trend.


